---
title: "Diamond Exploration"
output: html_document
---
### Included Libraries and Data Summary
```{r Libraries & Dataset}
library(ggplot2)
library(plyr)
library(dplyr)
data(diamonds)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

# Problem Set - Explore One Variables

### 1. Diamonds
> There are 53940 observations in data set

> It contains 10 variables.

> 3 ordered factors

> Best color of a diamond - 'D'

```{r Diamonds}
summary(diamonds)
```

### 2. Price Histogram
```{r Price Histogram}
qplot(x= price, data = diamonds)
```

### 3. Price Histogram Summary
> The shape of the graph is decreasing logarmithic.

> The Median price is 2401.

> The Mean price is 3933.

```{r}
summary(diamonds$price)
```

### 4. Diamond Counts
> 1729 diamonds cost less than $500.

> 0 diamonds cost less than $500.

> 1656 diamonds cost $15,000 or more.

```{r}
summary(diamonds$price < 500)
summary(diamonds$price > 250)
summary(diamonds$price >= 15000)
```


### 5. Cheaper Diamonds
```{r Cheaper Diamonds}
qplot(price, data =diamonds, binwidth = 10)+
  scale_x_continuous(limits= c(250,1000))
```

### 6. Price by Cut Histograms
```{r Price by Cut Histograms}
qplot(x= price, data = diamonds)+
  facet_wrap(~cut, ncol = 3)
```

### 7. Price by Cut
> 'Premium' cut has the highest priced diamond.

> 'Premium & Ideal' cut has the lowest priced diamond.

> 'Ideal' cut has the lowest median price.

```{r}
by(diamonds$price, diamonds$cut, summary)
```

### 8. Scales and Multiple Histograms
```{r Scales and Multiple Histograms}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales="free_y")
```


### 9.Price per Carat by Cut
```{r Price per Carat by Cut}
qplot(x = price, y= carat, data = diamonds) +
  facet_wrap(~cut, scales="free_y")
```

### 10. Price Box Plots
```{r Price Box Plots}
qplot(y=price/carat, x=color,
      data=na.omit(diamonds), geom = 'boxplot',
      fill = color)

qplot(y=price/carat, x=cut,
      data=na.omit(diamonds), geom = 'boxplot',
      fill = cut)

qplot(y=price/carat, x=clarity,
      data=na.omit(diamonds), geom = 'boxplot',
      fill = clarity)
```

### 11. Interquartile Range - IQR
> Price Range for middle 50% for color 'D' is *911* to *4214*.

> Price Range for middle 50% for color 'J' is *1860* to *7695*.

> IQR for best color is *3302.5*

> IQR for best color is *5834.5*
```{r Interquartile Range - IQR}
by(diamonds$price,diamonds$color,summary)
by(diamonds$price,diamonds$color,IQR)
```


### 12. Price per Carat Box Plots by Color
```{r Price per Carat Box Plots by Color}
qplot(y=price/carat, x=color,
      data=na.omit(diamonds), geom = 'boxplot',
      fill = color)
```

### 13. Carat Frequency Polygon
> Carat sizes having a count greater than 2000 are *0.3* & *1.01*

```{r Carat Frequency Polygon}
qplot(x=carat,data=na.omit(diamonds),
      geom = 'freqpoly',binwidth = 0.1,yaxs="i") + 
  scale_y_continuous(expand = c(0, 0),
                     breaks = seq(0,11000,1000),limits = c(0, 11000)) + 
  scale_x_continuous(expand = c(0, 0),
                     breaks = seq(0,5,0.1), limits = c(0, 5))
```


# Problem Set - Explore Two Variables

### 1. Price vs x
```{r Price vs x}
ggplot(aes(x,price), data = diamonds) + 
  geom_point()
```

### 2. Findings - price vs x
> Price rises exponentially for x b/w 6 to 9***

### 3. Correlations
```{r Correlations}
cor.test(diamonds$x,diamonds$price)
cor.test(diamonds$y,diamonds$price)
cor.test(diamonds$z,diamonds$price)
```

### 4. price vs depth
```{r price vs depth}
ggplot(aes(depth,price), data = diamonds) + 
  geom_point()
```

### 5. Adjustments- price vs depth
```{r Adjustments- price vs depth}
summary(diamonds$depth)

ggplot(aes(depth,price), data = diamonds) + 
  geom_point(alpha = 1/100, position = position_jitter(h=0))+
  scale_x_continuous(breaks = seq(43,79,2))
```

### 6. Typical Depth range
> 59 to 63

### 7. Correlation- price vs depth
```{r Correlation- price vs depth}
cor.test(diamonds$depth,diamonds$price)
```

### 8. price vs carat
```{r price vs carat}
ggplot(aes(carat,price), data = diamonds)+
  geom_point()+
  xlim(0, quantile(diamonds$carat, 0.99))+
  ylim(0, quantile(diamonds$price, 0.99))
```

### 9. price vs volume
```{r price vs volume}
diamonds$volume <- diamonds$x*diamonds$y*diamonds$z
summary(diamonds$volume)
ggplot(aes(volume,price), data = diamonds)+
  geom_point()+
  xlim(0,450)
```

### 10. Findings- price vs volume
> Both vary somewhat linerly

```{r Findings- price vs volume}
#count(diamonds$volume == 0)
```

### 11. Coorelations on subsets
```{r Coorelations on subsets}
with(subset(diamonds, volume>0 & volume<=800), cor.test(volume, price))
```

### 12. Adjustments - price vs volume
```{r Adjustments - price vs volume}
ggplot(aes(volume,price),data = subset(diamonds, volume>0 & volume<=800))+
  geom_point(alpha = 1/50, position = position_jitter(h=0))+
  geom_smooth(method = 'lm', color = 'red')+
  xlim(0,450)
```

### 13. Mean Price by clarity
```{r Mean Price by clarity}
clarityGroups <- group_by(diamonds, clarity)
diamondsByClarity <- summarize(
  clarityGroups,
  mean_price = mean(price),
  median_price = median(price),
  min_price = min(price),
  max_price = max(price),
  n = n())

head(diamondsByClarity)
```

### 14. Bar charts of mean price
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

library(gridExtra)

p1 <-ggplot(aes(clarity, mean_price), data=diamonds_mp_by_clarity)+
  geom_bar(stat = 'identity')

p2 <-ggplot(aes(color, mean_price), data=diamonds_mp_by_color)+
  geom_bar(stat = 'identity')

grid.arrange(p1, p2)

```


### 15. Trends in mean price
> SI2 has max of mean_price in Clarity.

> Mean_price increases form D to J in color.


# Problem Set - Explore Many Variables

### 1. Price Histograms with Facet and color
```{r Price Histograms with Facet and color}
qplot(data = diamonds, x= log(price), fill=cut) +
  facet_wrap(~color, ncol = 3) +
  ylim(0,600) +
  scale_fill_brewer(type = 'qual')
```

### 2. Price vs Table colored by cut
```{r Price vs Table colored by cut}
ggplot(aes(table,price),data = diamonds) +
  geom_point(aes(colour=cut))+
  scale_x_continuous(limits = c(50, 80), breaks = seq(50,80,2))+
  scale_fill_brewer(type = 'qual')
```

### 3. Typical Table Value
> Typical range for Ideal cut is 53 to 57.

> Typical range for Premium cut is 58 to 62.

### 4. Price vs. Volume and Diamond Clarity
```{r Price vs. Volume and Diamond Clarity}
ggplot(aes(volume, price), data = diamonds) +
  geom_point(aes(colour=clarity))+
  scale_fill_brewer(type = 'div')+
  scale_y_log10()+
  xlim(0, quantile(diamonds$volume, 0.99))
```

### 5. Proportion of Friendships Initiated
```{r Proportion of Friendships Initiated}
pf$prop_initiated <-pf$friendships_initiated/pf$friend_count
```

### 6. prop_initiated vs. tenure
```{r prop_initiated vs. tenure} 
pf$year_joined <- floor((2014 - pf$tenure/365))
pf$year_joined.bucket <- cut(
  pf$year_joined, breaks = c(2004,2009,2011,2012,2014))

table(pf$year_joined.bucket)
ggplot(aes(tenure, prop_initiated),
       data =subset(pf,!is.na(prop_initiated),!is.na(prop_initiated)))+
  geom_line(aes(colour= year_joined.bucket), stat = 'summary', fun.y = median)
```

### 7. Smoothing prop_initiated vs. tenure
```{r Smoothing prop_initiated vs. tenure}
ggplot(aes(tenure, prop_initiated),
       data =subset(pf,!is.na(prop_initiated),!is.na(prop_initiated)))+
  geom_smooth(aes(colour= year_joined.bucket))

```

### 8. Greatest prop_initiated Group
> People who joined after 2012, initiated the greatest proportion of its freindships.

### 9. Largest Group Mean prop_initiated
> Mean of Largest Group = 0.6654

> The reason behind the enormous increase in friendships initiated after 2012, is the rapid growing popularity of facebook among the younger age groups, who spend a lot of time on suck social networking sites. 

```{r}
by(pf$prop_initiated, pf$year_joined.bucket, summary)
```

### 10. Price/Carat Binned, Faceted, & Colored
```{r Price/Carat Binned, Faceted, & Colored}
ggplot(aes(cut,price/carat), data= diamonds)+
  geom_jitter(aes(colour=color))+
  facet_wrap(~clarity, ncol = 3)+
  scale_color_brewer(type = 'div')+
  theme(axis.text=element_text(size=6),
        axis.title=element_text(size=14,face="bold"))
```


